---
title: "FFSFuels EDA"
author: "Jacob_Levine"
date: "9/12/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fire Fire Surrogate Fuels Accumulation - Exploratory Data Analysis

The purpose of this file is to develop figures and statistics in order to better understand accumulation of fine and woody fuels in the Fire Fire Surrogate Study Area.


## Load required packages
```{r}
library(here)
library(reshape2)
library(ggplot2)
library(gridExtra)
library(nlme)
```

#### Run data-cleaning file:
```{r}

source("FFSFuels_DataClean.R")

```

## Look at distribution of fuel loads:

```{r}
## plot histograms:

## Impose PDF on straight and log-transformed data:
## function to compute summary stats
sumstatdf <- function() {
  sumstat <- data.frame(FuelClass = c("total", "litter", "duff", "litterduff", "1h", "10h", "100h", "1000s", "1000r", "1.100h", "1.1000h", "surface"))

  meantotal <- mean(FFS_data$fuelload_total_tonha)
  vartotal <- var(FFS_data$fuelload_total_tonha)
  logmeantotal <- mean(FFS_data$logfuelload_total)
  logvartotal <- var(FFS_data$logfuelload_total)
  
  meanlitter <- mean(FFS_data$fuelload_litter_tonha)
  varlitter <- var(FFS_data$fuelload_litter_tonha)
  logmeanlitter <- mean(FFS_data$logfuelload_litter)
  logvarlitter <- var(FFS_data$logfuelload_litter)
  
  meanduff <- mean(FFS_data$fuelload_duff_tonha)
  varduff <- var(FFS_data$fuelload_duff_tonha)
  logmeanduff <- mean(FFS_data$logfuelload_duff)
  logvarduff <- var(FFS_data$logfuelload_duff)
  
  meanlitterduff <- mean(FFS_data$fuelload_litterduff_tonha)
  varlitterduff <- var(FFS_data$fuelload_litterduff_tonha)
  logmeanlitterduff <- mean(FFS_data$logfuelload_litterduff)
  logvarlitterduff <- var(FFS_data$logfuelload_litterduff)
  
  mean1h <- mean(FFS_data$fuelload_1h_tonha)
  var1h <- var(FFS_data$fuelload_1h_tonha)
  logmean1h <- mean(FFS_data$logfuelload_1h)
  logvar1h <- var(FFS_data$logfuelload_1h)
  
  mean10h <- mean(FFS_data$fuelload_10h_tonha)
  var10h <- var(FFS_data$fuelload_10h_tonha)
  logmean10h <- mean(FFS_data$logfuelload_10h)
  logvar10h <- var(FFS_data$logfuelload_10h)
  
  mean100h <- mean(FFS_data$fuelload_100h_tonha)
  var100h <- var(FFS_data$fuelload_100h_tonha)
  logmean100h <- mean(FFS_data$logfuelload_100h)
  logvar100h <- var(FFS_data$logfuelload_100h)
  
  mean1000s <- mean(FFS_data$fuelload_1000s_tonha)
  var1000s <- var(FFS_data$fuelload_1000s_tonha)
  logmean1000s <- mean(FFS_data$logfuelload_1000s)
  logvar1000s <- var(FFS_data$logfuelload_1000s)
  
  mean1000r <- mean(FFS_data$fuelload_1000r_tonha)
  var1000r <- var(FFS_data$fuelload_1000r_tonha)
  logmean1000r <- mean(FFS_data$logfuelload_1000r)
  logvar1000r <- var(FFS_data$logfuelload_1000r)
  
  mean1.100h <- mean(FFS_data$fuelload_1.100h_tonha)
  var1.100h <- var(FFS_data$fuelload_1.100h_tonha)
  logmean1.100h <- mean(FFS_data$logfuelload_1.100h)
  logvar1.100h <- var(FFS_data$logfuelload_1.100h)
  
  mean1.1000h <- mean(FFS_data$fuelload_1.1000h_tonha)
  var1.1000h <- var(FFS_data$fuelload_1.1000h_tonha)
  logmean1.1000h <- mean(FFS_data$logfuelload_1.1000h)
  logvar1.1000h <- var(FFS_data$logfuelload_1.1000h)
  
  meansurface <- mean(FFS_data$fuelload_surface_tonha)
  varsurface <- var(FFS_data$fuelload_surface_tonha)
  logmeansurface <- mean(FFS_data$logfuelload_surface)
  logvarsurface <- var(FFS_data$logfuelload_surface)

  sumstat$mean <- c(meantotal, meanlitter, meanduff, meanlitterduff, mean1h, mean10h, mean100h, mean1000s, mean1000r, mean1.100h, mean1.1000h, meansurface)
  sumstat$var <- c(vartotal, varlitter, varduff, varlitterduff, var1h, var10h, var100h, var1000s, var1000r, var1.100h, var1.1000h, varsurface)
  sumstat$logmean <- c(logmeantotal, logmeanlitter, logmeanduff, logmeanlitterduff, logmean1h, logmean10h, logmean100h, logmean1000s, logmean1000r, logmean1.100h, logmean1.1000h, logmeansurface)
  sumstat$logvar <- c(logvartotal, logvarlitter, logvarduff, logvarlitterduff, logvar1h, logvar10h, logvar100h, logvar1000s, logvar1000r, logvar1.100h, logvar1.1000h, logvarsurface)
  return(sumstat)
}
## write function to dataframe
sumstat <- sumstatdf()

## function to generate plots
makereghists <- function(row, stats = sumstat, data = FFS_data) {
    ## create key variable names:
    reg <- paste0("fuelload_", row, "_tonha")
    ## create sequences for regular and log:
    seqvalreg <- seq(min(data[, reg]), max(data[, reg]), length.out = nrow(data))
    binwid <- (max(data[, reg]) - min(data[, reg]))/25 
    
    ## create graphs:
    ggplot(data = data, aes(data[,reg])) +
                  geom_histogram(aes(y = binwid*..density..), fill = "steelblue3", binwidth = binwid) + 
                  geom_line(aes(x = seqvalreg, y = binwid*dnorm(x = seqvalreg, mean = stats[stats$FuelClass == row, "mean"], sd = sqrt(stats[stats$FuelClass == row, "var"])))) +
      xlab(reg) +
      ylab("20 * Relative Frequency")
}

makeloghists <- function(row, stats = sumstat, data = FFS_data) {
    ## create key variable names:
    log <- paste0("logfuelload_", row)
    ## create sequences for regular and log:
    seqvallog <- seq(min(data[, log]), max(data[, log]), length.out = nrow(data))
    binwid <- (max(data[, log]) - min(data[, log]))/25
    
    ## create graph:
    ggplot(data = data, aes(data[,log])) +
                  geom_histogram(aes(y = binwid*..density..), fill = "steelblue3", binwidth = binwid) + 
                  geom_line(aes(x = seqvallog, y = binwid*dnorm(x = seqvallog, mean = stats[stats$FuelClass == row, "logmean"], sd = sqrt(stats[stats$FuelClass == row, "logvar"])))) +
                  xlab(log) +
                  ylab("20 * Relative Frequency") 
}

regplots <- lapply(sumstat$FuelClass, makereghists, data = FFS_data)
regplots

logplots <- lapply(sumstat$FuelClass, makeloghists, data = FFS_data)
logplots

## lets check distribution by treatment:

ggplot(data = FFS_data, aes(fuelload_total_tonha)) +
  geom_histogram(fill = "steelblue3", bins = 25) +
  facet_grid(~Treatment)

## check distribution of litter&duff vs fine vs coarse fuels:

ggplot(data = FFS_data, aes(fuelload_total_tonha)) +
  geom_histogram(fill = "steelblue3", bins = 25) 


```


Appears to follow a Poisson distribution (roughly)

## Plot development of fuels chronologically

## determine average difference in date between Pretreatment and Post_1 measurements:
```{r}

#### subset FFS data
dates_pre1 <- subset(FFS_data, FFS_data$Timestep == "Pretreatment" | FFS_data$Timestep == "Post_1")
dates_pre1_fuels <- dates_pre1[, c("Invdate.fuels", "PlotID", "Timestep")]
dates_pre1_trees <- dates_pre1[, c("Invdate.trees", "PlotID", "Timestep")]

#### pivot tables
dates_pre1_fuels <- dcast(dates_pre1_fuels, PlotID ~ Timestep, value.var = "Invdate.fuels")
colnames(dates_pre1_fuels) <- c("PlotID", "Post_1Fuels", "PretreatmentFuels")

dates_pre1_trees <- dcast(dates_pre1_trees, PlotID ~ Timestep, value.var = "Invdate.trees")
colnames(dates_pre1_trees) <- c("PlotID", "Post_1Trees", "PretreatmentTrees")

#### combine tables
dates_pre1 <- merge.data.frame(dates_pre1_trees, dates_pre1_fuels)

#### coerce factor to date
dates_pre1$Post_1Trees <- as.Date(dates_pre1$Post_1Trees)
dates_pre1$PretreatmentTrees <- as.Date(dates_pre1$PretreatmentTrees)
dates_pre1$Post_1Fuels <- as.Date(dates_pre1$Post_1Fuels)
dates_pre1$PretreatmentFuels <- as.Date(dates_pre1$PretreatmentFuels)

#### get time between measurements for fuels and trees
dates_pre1$diffTrees <- dates_pre1$Post_1Trees - dates_pre1$PretreatmentTrees
dates_pre1$diffFuels <- dates_pre1$Post_1Fuels - dates_pre1$PretreatmentFuels

meandiffFuels <- mean(dates_pre1$diffFuels)
meandiffTrees <- mean(dates_pre1$diffTrees)

#### Difference between measurements roughly 2 years, implying treatment occurs roughly 1 year after pretreatment 

## plot total fuel by timestep
ggplot(data = FFS_data, aes(x = FFS_data$TimestepNum, y = (FFS_data$fuelload_total_tonha))) +
  geom_jitter(aes(color = Treatment)) +
  geom_smooth(aes(color = Treatment)) 

ggplot(data = FFS_data, aes(x = FFS_data$TimestepNum, y = (FFS_data$fuelload_total_tonha))) +
  geom_point(aes(color = Treatment)) +
  geom_smooth(aes(color = Treatment))

## plot total fuel by timestep as categorical var.
stocktimetotal <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelload_total_tonha)) +
  geom_boxplot(aes(color = Treatment)) +
  labs(x = "Timestep", y = "Total Fuels (ton/ha)")

## for litterduff
stocktimelitterduff <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelload_litterduff_tonha)) +
  geom_boxplot(aes(color = Treatment)) +
  labs(x = "Timestep", y = "Litter and Duff (ton/ha)")

## for 1.100h
stocktime1.100h <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelload_1.100h_tonha)) +
  geom_boxplot(aes(color = Treatment)) +
  labs(x = "Timestep", y = "Litter and Duff (ton/ha")
```


# Deltas as response variable:

#### graph deltas:

###### Histograms
```{r}
FFS_data_complete <- subset(FFS_data, !is.na(FFS_data$fuelchange_total))

## plot overall distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_total)) +
  geom_histogram(fill = "steelblue3", bins = 30) 

## plot litterduff distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_litterduff)) +
  geom_histogram(fill = "steelblue3", bins = 30) 

## plot 1.100 distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_1.100h)) +
  geom_histogram(fill = "steelblue3", bins = 30)

## look at same series, by treatment:
## plot overall distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_total)) +
  geom_histogram(fill = "steelblue3", bins = 20) +
  facet_grid(Timestep~Treatment)

## plot litterduff distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_litterduff)) +
  geom_histogram(fill = "steelblue3", bins = 30) +
  facet_grid(Timestep~Treatment)

## plot 1.100 distribution:
ggplot(data = FFS_data_complete, aes(fuelchange_1.100h)) +
  geom_histogram(fill = "steelblue3", bins = 30) +
  facet_grid(Timestep~Treatment)



```

###### By density:

```{r}

## without nesting:
ggplot(data = FFS_data_complete, aes(x = BA.mha.sum_Live_.all., y = fuelchange_total)) +
  geom_jitter() +
  geom_smooth(method = lm) 


ggplot(data = FFS_data_complete, aes(x = BA.mha.sum_Live_.all., y = fuelchange_total)) +
  geom_jitter() +
  geom_smooth(method = lm) + 
  facet_grid(Treatment~Timestep)

```

###### Time Series
```{r}

deltatimetotal <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelchange_total)) +
  geom_boxplot(aes(color = Treatment)) +
  labs(x = "", y = "Change in Total Fuels (ton/ha)") 
  

deltatimelitterduff <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelchange_litterduff)) +
  geom_boxplot(aes(color = Treatment)) +
  labs(x = "", y = "Change in Litter and Duff (ton/ha)") 

deltatime1.100h <- ggplot(data = FFS_data, aes(x = FFS_data$Timestep, y = FFS_data$fuelchange_1.100h)) +
  geom_boxplot(aes(color = Treatment)) + 
  labs(x = "", y = "Change in Fine fuels (1-100h, ton/ha)") 

```


#### compare rates to stocks
```{r}
### for total fuels
grid.arrange(deltatimetotal, stocktimetotal, ncol = 1)

### for litterduff
grid.arrange(deltatimelitterduff, stocktimelitterduff, ncol = 1)

### for 1.100h
grid.arrange(deltatime1.100h, stocktime1.100h, ncol = 1)

```

Note that control fine fuels reduce significantly between Post_1 and Post_7
this doesn't make a lot of ecological sense - perhaps a crew effect? 


## Calculate rates of change:

```{r}

```

## graph rates:

```{r}
FFS_data_complete <- subset(FFS_data, !is.na(FFS_data$fuelrate_total))


## remove outliers

FFS_data_complete <- subset(FFS_data_complete, FFS_data_complete$fuelrate_total < 10)

ggplot(data = FFS_data_complete, aes(x = FFS_data_complete$Timestep, y = FFS_data_complete$fuelrate_total)) +
  geom_boxplot(aes(color = Treatment))

```


Examine difference between compartments:

```{r}

## put compartments in order of treatment:
FFS_data$Unit <- as.factor(FFS_data$Unit)
FFS_data$Unit <- factor(FFS_data$Unit, levels = c("40", "240", "590", "60", "340", "400", "190", "350", "490", "180", "380", "570"))

## create dummy variable col for purpose of graphing replicates
FFS_data$TreatReplicate <- factor(rep(NA, times = nrow(FFS_data)), levels = c("1", "2", "3"))

## fill in replicate column:
for (i in 1:4) {
  curdf <- FFS_data[FFS_data$Treatment == levels(FFS_data$Treatment)[i], ]
  curUnits <- unique(curdf$Unit)
  for (j in 1:3) {
    FFS_data[FFS_data$Treatment == levels(FFS_data$Treatment)[i] &
               FFS_data$Unit == curUnits[j], "TreatReplicate"] <- as.factor(j)
  }
}

## put treatments in right order:
FFS_data$Treatment <- factor(FFS_data$Treatment, levels = c("Control", "Burn", "Mech", "MechBurn"))

## plot replicate by treatment
ggplot(data = FFS_data, aes(x = Timestep, y = fuelload_total_tonha)) + 
  geom_boxplot(aes(color = Treatment)) + 
  facet_grid(TreatReplicate~Treatment) +
  theme(legend.position = "none") + 
  ylab(c("Total Fuel Load (tonHa)", "Replicate"))

```

## Examine difference according to species composition:

```{r}

## create columns for species comp ratios
FFS_data$SPPRatioPIPO <- FFS_data$BA.mha.sum_Live_PIPO/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioPILA <- FFS_data$BA.mha.sum_Live_PILA/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioPSME <- FFS_data$BA.mha.sum_Live_PSME/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioQUKE <- FFS_data$BA.mha.sum_Live_QUKE/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioABCO <- FFS_data$BA.mha.sum_Live_ABCO/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioCADE <- FFS_data$BA.mha.sum_Live_CADE/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioLIDE <- FFS_data$BA.mha.sum_Live_LIDE/FFS_data$BA.mha.sum_Live_.all.

FFS_data$SPPRatioOTHER <- FFS_data$BA.mha.sum_Live_OTHER/FFS_data$BA.mha.sum_Live_.all.

## subset to look only at pretreatment data:

presub_data <- subset(FFS_data, FFS_data$Timestep == "Post_14")

FFS_data$logFUELTotal <- log(FFS_data$fuelload_total_tonha)


ggplot(data = FFS_data, aes(x = BA.mha.sum_Live_.all., y = fuelload_total_tonha)) +
  geom_jitter() +
  geom_smooth(method = lm) + 
  facet_grid(Treatment~Timestep)
  
```






