---
title: "FFSFuels_BurnPlots"
author: "Jacob_Levine"
date: "9/25/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# FFSFuels Burn Treatment - Fuel Accumulation

#### Load required packages
```{r}
library(pscl)
library(MuMIn)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(nlme)
library(lme4)
```

#### Set up data:

Load in data
```{r}
## load in data
path <- here("/Fuels_byplot_Plots_ffs_filtered_postburns.csv") 
FFS_data <- read.csv(path, header = TRUE)
## extract burn plots
fire_data <- subset(FFS_data, Treatment == "Burn")
```


Calculate consumption for total fuelloads, duff, litter, 1h, 10h, 100h, 1000h.
```{r}
## calculate consumption for total fuelloads for now (after burn 1 and burn 2, not sure about burn 3 yet)
## expressed as a percent

calc_cons <- function(data, fuels) {
  
  ## add columns for Unit, PlotID and BurnNumber
  Unit <- rep(data[data$Timestep == "Pretreatment" | data$Timestep == "Post_7" | data$Timestep == "Post_14", "Unit"], times = length(fuels))
  PlotID <- rep(data[data$Timestep == "Pretreatment" | data$Timestep == "Post_7" | data$Timestep == "Post_14", "PlotID"])
  d <- expand.grid(PlotID, fuels)
  colnames(d) <- c("PlotID", "FuelClass")
  BurnID <- rep(c(1,2,3), times = nrow(d)/3)
  
  df <- data.frame(Unit, d, BurnID)
  
  ## calculate consumption for each burnnumber, fuel category, and plot
    for (j in 1:length(fuels)) {
      fuel <- paste0("fuelload_", fuels[j], "_tonha")
      for (k in 1:length(unique(data$PlotID))) {
         for (i in 1:3) {
          ## set correct timestep based on BurnNumber
          if (i == 1) {
            timestepPre <- "Pretreatment"
            timestepPost <- "Post_burn1"
          }
          else if (i == 2) {
            timestepPre <- "Post_7"
            timestepPost <- "Post_burn2"
          }
          else if (i == 3) {
            timestepPre <- "Post_14"
            timestepPost <- "Post_burn3"
          }
          cons <- (-1 * (data[data$PlotID == unique(data$PlotID)[k] &
                         data$Timestep == timestepPost, fuel] - data[data$PlotID == unique(data$PlotID)[k] &
                         data$Timestep == timestepPre, fuel]) / data[data$PlotID == unique(data$PlotID)[k] &
                         data$Timestep == timestepPre, fuel] * 100)
          consNet <- (-1 * (data[data$PlotID == unique(data$PlotID)[k] &
                         data$Timestep == timestepPost, fuel] - data[data$PlotID == unique(data$PlotID)[k] &
                         data$Timestep == timestepPre, fuel]))
        
          df[df$PlotID == unique(data$PlotID)[k] &
             df$BurnID == i &
             df$FuelClass == fuels[j], "Consumption"] <- cons
          
          df[df$PlotID == unique(data$PlotID)[k] &
             df$BurnID == i &
             df$FuelClass == fuels[j], "NetConsumption"] <- consNet
          
          
          }
        }
      }
      return(df)
    }

fuelslist <- c("total", "duff", "litter", "1h", "10h", "100h", "1000h")
cons_data <- calc_cons(data = fire_data, fuels = fuelslist)

```

## Consumption

Questions to answer:

1) Quantify consumption of fuels
    -group by total, litter, duff, 1, 10, 100, 1000.
    -look at difference in consumption from burn 1 to burn 2.
    -create table and graphs that examine this difference.
    -Look at consumption in mechburn vs burn plots.
    -variablity in consumption between plots (spatial heterogeneity?)

2) What drives the difference in consumption between plots.
    - Model selection (using all types of fuelload as explanatory variables?  Unit as random variable?)
    - 2 part process:  did the plot burn at all?  If so: how much did it burn.


#### Plotting consumption
```{r}
  


## remove data wherein fuel level increased after fire and NAs
cons_data1 <- cons_datapiv[cons_datapiv$total > 0 & cons_datapiv$total < 200, ]
cons_data1 <- cons_data1[complete.cases(cons_data1$total),]
cons_data <- cons_data[cons_data$Consumption > 0 & cons_data$NetConsumption < 200,]
cons_data <- cons_data[complete.cases(cons_data$Consumption),]

## plotting change
ggplot(cons_data, aes(y = Consumption, x = as.factor(FuelClass), fill = as.factor(BurnID))) +
  geom_boxplot() +
  guides(fill = guide_legend(title = "Burn Number")) +
  xlab("Fuel Class") + 
  ylab("Consumption (% of pre burn)")


ggplot(subset(cons_data, cons_data$FuelClass != "total"), aes(y = Consumption, x = as.factor(FuelClass), fill = as.factor(BurnID))) +
  geom_boxplot() +
  guides(fill = guide_legend(title = "Burn Number")) +
  facet_grid(Unit~.) +
  xlab("Fuel Class") + 
  ylab("Consumption (% of pre burn)")


ggplot(cons_data, aes(y = NetConsumption, x = as.factor(FuelClass), fill = as.factor(BurnID))) +
  geom_boxplot() +
  guides(fill = guide_legend(title = "Burn Number")) +
  xlab("Fuel Class") + 
  ylab("Consumption (ton/ha)")


ggplot(subset(cons_data, cons_data$FuelClass != "total"), aes(y = NetConsumption, x = as.factor(FuelClass), fill = as.factor(BurnID))) +
  geom_boxplot() +
  guides(fill = guide_legend(title = "Burn Number")) +
  facet_grid(Unit~.) +
  xlab("Fuel Class") + 
  ylab("Consumption (% of pre burn)")


cons_data1$logtotal <- log(cons_data1$total)
cons_data1$logloadtotal <- log(cons_data1$fuelload_total_tonha)

tt <- ggplot(cons_data1, aes(y = total, x = fuelload_total_tonha)) +
  geom_point() +
  geom_smooth() +
  ylab("Total Consumption")

td <- ggplot(cons_data1, aes(y = duff, x = fuelload_duff_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("Sudd Consumption")

tl <- ggplot(cons_data1, aes(y = litter, x = fuelload_litter_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("Litter Consumption")

t1 <- ggplot(cons_data1, aes(y = X1h, x = fuelload_1h_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("1h Consumption")

t10 <- ggplot(cons_data1, aes(y = X10h, x = fuelload_10h_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("10h Consumption")

t100 <- ggplot(cons_data1, aes(y = X100h, x = fuelload_100h_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("100h Consumption")

t1000 <- ggplot(cons_data1, aes(y = X1000h, x = fuelload_1000h_tonha)) +
  geom_point() +
  geom_smooth()+
  ylab("1000h Consumption")

grid.arrange(tt, td, tl, t1, t10, t100, t1000)

```


## Model Fitting for Consumption

#### Data Prep

```{r}

## calculate species ratios
fire_data$SPPRatioPIPO <- fire_data$BA.mha.sum_Live_PIPO/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioPILA <- fire_data$BA.mha.sum_Live_PILA/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioPSME <- fire_data$BA.mha.sum_Live_PSME/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioQUKE <- fire_data$BA.mha.sum_Live_QUKE/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioABCO <- fire_data$BA.mha.sum_Live_ABCO/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioCADE <- fire_data$BA.mha.sum_Live_CADE/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioLIDE <- fire_data$BA.mha.sum_Live_LIDE/fire_data$BA.mha.sum_Live_.all. 
fire_data$SPPRatioOTHER <- fire_data$BA.mha.sum_Live_OTHER/fire_data$BA.mha.sum_Live_.all. 

## Pivot for Consumption
cons_datapiv <- dcast(cons_data[1:5], Unit + PlotID + BurnID ~ FuelClass)
firesub <- subset(fire_data, fire_data$Timestep == "Pretreatment" | fire_data$Timestep == "Post_7" | fire_data$Timestep == "Post_14")
cons_datapiv <- data.frame(cons_datapiv, firesub[, c("BA.mha.sum_Live_.all.", "BA.mha.sum_Snag_.all.", "SPPRatioPIPO", "SPPRatioPILA", "SPPRatioPSME", "SPPRatioQUKE", "SPPRatioABCO", "SPPRatioCADE", "SPPRatioLIDE", "SPPRatioOTHER", "fuelload_total_tonha", "fuelload_litter_tonha", "fuelload_duff_tonha", "fuelload_1h_tonha", "fuelload_10h_tonha", "fuelload_100h_tonha", "fuelload_1000h_tonha")])

## pivot for NetConsumption
cons_datapiv1 <- dcast(cons_data[,c(1:4,6)], Unit + PlotID + BurnID ~ FuelClass)
firesub <- subset(fire_data, fire_data$Timestep == "Pretreatment" | fire_data$Timestep == "Post_7" | fire_data$Timestep == "Post_14")

## put Netconsumption into cons_datapiv
cons_datapiv$totalNet <- cons_datapiv1$total
cons_datapiv$duffNet <- cons_datapiv1$duff
cons_datapiv$litterNet <- cons_datapiv1$litter
cons_datapiv$X1hNet <- cons_datapiv1$X1h
cons_datapiv$X10hNet <- cons_datapiv1$X10h
cons_datapiv$X100hNet <- cons_datapiv1$X100h
cons_datapiv$X1000hNet <- cons_datapiv1$X1000h

cons_datapiv$logfuelload_total <- log(cons_datapiv$fuelload_total_tonha)

## rename for ease.
Cdata <- cons_datapiv

```

#### For just Burn 1:

```{r}
CdataB1 <- subset(Cdata, BurnID == "1")

lmeB1 <- lme(data = CdataB1, total ~ BA.mha.sum_Live_.all. + BA.mha.sum_Snag_.all. + SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME, random = ~1|Unit)

dredge(lmeB1)

### best fitting model:
lmeB1_1 <- lme(data = CdataB1, total ~ BA.mha.sum_Live_.all. + SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)
summary(lmeB1_1)
anova(lmeB1_1)

```

#### For just Burn 2:
```{r}
CdataB2 <- subset(Cdata, BurnID == "2")

lmeB2 <- lme(data = CdataB2, total ~ BA.mha.sum_Live_.all. + BA.mha.sum_Snag_.all. + SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME, random = ~1|Unit)

dredge(lmeB2)

### best fitting model:
lmeB2_1 <- lme(data = CdataB2, total ~ SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)
summary(lmeB2_1)
anova(lmeB2_1)

```

#### For just Burn 3:

```{r}
CdataB3 <- subset(Cdata, BurnID == "3")

lmeB3 <- lme(data = CdataB3, total ~ BA.mha.sum_Live_.all. + BA.mha.sum_Snag_.all. + SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)

dredge(lmeB3)

lmeB3_1 <- lme(data = CdataB3, total ~ BA.mha.sum_Snag_.all. + SPPRatioABCO + SPPRatioCADE  + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)
summary(lmeB3_1)
anova(lmeB3_1)
```

#### For all Burns together

```{r}
lmeT <- lme(data = Cdata, total ~ BA.mha.sum_Live_.all. + BA.mha.sum_Snag_.all. + SPPRatioABCO + SPPRatioCADE + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)

dredge(lmeT)

lmeBT_1 <- lme(data = Cdata, total ~ SPPRatioABCO + SPPRatioCADE  + SPPRatioPILA + SPPRatioPIPO + SPPRatioPSME + logfuelload_total, random = ~1|Unit)
summary(lmeBT_1)
anova(lmeBT_1)
```



```{r}

ggplot(Cdata, aes(SPPRatioPILA)) +
  geom_histogram()

```






















## What effects fuel accumulation between post_1 and post_7?

#### data preparation

```{r}
## pull out data:
fire_data1_7 <- subset(fire_data, fire_data$Timestep == "Post_1" | fire_data$Timestep == "Post_7")

calc_deltas1_7 <- function(data = fire_data1_7) {
  ## initialize vectors
  fuelchangeBACI_litter <- numeric()
  fuelchangeBACI_duff <- numeric()
  fuelchangeBACI_1000h <- numeric()
  fuelchangeBACI_1.100h <- numeric()
  fuelchangeBACI_total <- numeric()
  
  ## fill vectors with calculated values
  for (i in 1:length(unique(data$PlotID))) {
    for (j in 1:length(unique(data[data$PlotID == unique(data$PlotID)[i], "Timestep"]))) {
          ## litter
          NewfuelchangeBACI_litter <- ((data[data$PlotID == unique(data$PlotID)[i] &
                                              data$Timestep == "Post_7", "fuelload_litter_tonha"] - data[data$PlotID == unique(data$PlotID)[i] &
                                                                                                                      data$Timestep == "Post_1", "fuelload_litter_tonha"]))
          fuelchangeBACI_litter <- c(fuelchangeBACI_litter, NewfuelchangeBACI_litter)
          
          ## duff
          NewfuelchangeBACI_duff <- ((data[data$PlotID == unique(data$PlotID)[i] &
                                            data$Timestep == "Post_7", "fuelload_duff_tonha"] - data[data$PlotID == unique(data$PlotID)[i] &
                                                                                                                  data$Timestep == "Post_1", "fuelload_duff_tonha"]))
          fuelchangeBACI_duff <- c(fuelchangeBACI_duff, NewfuelchangeBACI_duff)
          
          ## 1000h
          NewfuelchangeBACI_1000h <- ((data[data$PlotID == unique(data$PlotID)[i] &
                                             data$Timestep == "Post_7", "fuelload_1000h_tonha"] - data[data$PlotID == unique(data$PlotID)[i] &
                                                                                                                    data$Timestep == "Post_1", "fuelload_1000r_tonha"]))
          fuelchangeBACI_1000h <- c(fuelchangeBACI_1000h, NewfuelchangeBACI_1000h)
          
          ## 1.100h
          NewfuelchangeBACI_1.100h <- ((data[data$PlotID == unique(data$PlotID)[i] &
                                              data$Timestep == "Post_7", "fuelload_1.100h_tonha"] - data[data$PlotID == unique(data$PlotID)[i] &
                                                                                                                      data$Timestep == "Post_1", "fuelload_1.100h_tonha"]))  
          fuelchangeBACI_1.100h <- c(fuelchangeBACI_1.100h, NewfuelchangeBACI_1.100h)
          
          ## total
          NewfuelchangeBACI_total <- ((data[data$PlotID == unique(data$PlotID)[i] &
                                             data$Timestep == "Post_7", "fuelload_total_tonha"] - data[data$PlotID == unique(data$PlotID)[i] &
                                                                                                                    data$Timestep == "Post_1", "fuelload_total_tonha"]))  
          fuelchangeBACI_total <- c(fuelchangeBACI_total, NewfuelchangeBACI_total)
          
    }
  }
  ## append new vectors to data:
  data$fuelchange_litter <- fuelchangeBACI_litter
  data$fuelchange_duff <- fuelchangeBACI_duff
  data$fuelchange_1000h <- fuelchangeBACI_1000h
  data$fuelchange_1.100h <- fuelchangeBACI_1.100h
  data$fuelchange_total <- fuelchangeBACI_total
  return(data)
}
fire_data1_7 <- calc_deltas1_7()

ggplot(subset(fire_data1_7, fire_data1_7$Timestep == "Post_1"), aes(fuelchange_total)) +
  geom_histogram(bins = 20)

fire_data1_7$SPPRatioPIPO <- fire_data1_7$BA.mha.sum_Live_PIPO/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioPILA <- fire_data1_7$BA.mha.sum_Live_PILA/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioPSME <- fire_data1_7$BA.mha.sum_Live_PSME/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioQUKE <- fire_data1_7$BA.mha.sum_Live_QUKE/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioABCO <- fire_data1_7$BA.mha.sum_Live_ABCO/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioCADE <- fire_data1_7$BA.mha.sum_Live_CADE/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioLIDE <- fire_data1_7$BA.mha.sum_Live_LIDE/fire_data1_7$BA.mha.sum_Live_.all. 
fire_data1_7$SPPRatioOTHER <- fire_data1_7$BA.mha.sum_Live_OTHER/fire_data1_7$BA.mha.sum_Live_.all. 

## calculate mean density by block:
blockmeans <- aggregate(x = fire_data1_7$BA.mha.sum_Live_.all., by = list(fire_data1_7$Unit), FUN = mean)

plotmeans <- aggregate(x = fire_data1_7$BA.mha.sum_Live_.all., by = list(fire_data1_7$Unit, fire_data1_7$PlotID), FUN = mean)

## append block means to fire_data1_7_14:
fire_data1_7$blockmean <- numeric(length = nrow(fire_data1_7))
fire_data1_7$BAdif <- numeric(length = nrow(fire_data1_7))
for (i in 1:nrow(blockmeans)) {
  fire_data1_7[fire_data1_7$Unit == blockmeans$Group.1[i], "blockmean"] <- blockmeans$x[i]
}

## so that BAdif is difference between average BA of plot for all times minus average BA of unit for all times:
for (i in 1:nrow(plotmeans)) {
  fire_data1_7[fire_data1_7$Unit == plotmeans$Group.1[i] &
                fire_data1_7$PlotID == plotmeans$Group.2[i], "BAdif"] <- plotmeans$x[i] - blockmeans[blockmeans$Group.1 == plotmeans$Group.1[i], "x"]
}



## calculate mean density by block:
blockmeansnag <- aggregate(x = fire_data1_7$BA.mha.sum_Snag_.all., by = list(fire_data1_7$Unit), FUN = mean)


plotmeansnag <- aggregate(x = fire_data1_7$BA.mha.sum_Snag_.all., by = list(fire_data1_7$Unit, fire_data1_7$PlotID), FUN = mean)

## append block means to fire_data1_7_14:
fire_data1_7$blockmeanSnag <- numeric(length = nrow(fire_data1_7))
fire_data1_7$BAdifSnag <- numeric(length = nrow(fire_data1_7))
for (i in 1:nrow(blockmeansnag)) {
  fire_data1_7[fire_data1_7$Unit == blockmeansnag$Group.1[i], "blockmeanSnag"] <- blockmeansnag$x[i]
}

## so that BAdif is difference between average BA of plot for all times minus average BA of unit for all times:
for (i in 1:nrow(plotmeansnag)) {
  fire_data1_7[fire_data1_7$Unit == plotmeansnag$Group.1[i] &
                fire_data1_7$PlotID == plotmeansnag$Group.2[i], "BAdifSnag"] <- plotmeansnag$x[i] - blockmeansnag[blockmeansnag$Group.1 == plotmeans$Group.1[i], "x"]
}


## calculate mean density by block:
blockmeancons <- aggregate(x = fire_data1_7$ConsumptionBurn1, by = list(fire_data1_7$Unit), FUN = mean)


plotmeancons <- aggregate(x = fire_data1_7$ConsumptionBurn1, by = list(fire_data1_7$Unit, fire_data1_7$PlotID), FUN = mean)

## append block means to fire_data1_7_14:
fire_data1_7$blockmeancons <- numeric(length = nrow(fire_data1_7))
fire_data1_7$BAdifcons <- numeric(length = nrow(fire_data1_7))
for (i in 1:nrow(blockmeancons)) {
  fire_data1_7[fire_data1_7$Unit == blockmeancons$Group.1[i], "blockmeancons"] <- blockmeancons$x[i]
}

## so that BAdif is difference between average BA of plot for all times minus average BA of unit for all times:
for (i in 1:nrow(plotmeancons)) {
  fire_data1_7[fire_data1_7$Unit == plotmeancons$Group.1[i] &
                fire_data1_7$PlotID == plotmeancons$Group.2[i], "difcons"] <- plotmeancons$x[i] - blockmeancons[blockmeancons$Group.1 == plotmeans$Group.1[i], "x"]
}



### Calculate mortality between Pretreatment and Post 7:
fire_data <- fire_data[complete.cases(fire_data$count.ha_Live_.all.),]
for (i in 1:length(unique(fire_data1_7$PlotID))) {
  fire_data1_7[fire_data1_7$PlotID == unique(fire_data1_7$PlotID)[i], "MortalityP_7"] <- fire_data[fire_data$PlotID == unique(fire_data1_7$PlotID)[i] & fire_data$Timestep == "Post_7", "count.ha_Live_.all."] - fire_data[fire_data$PlotID == unique(fire_data1_7$PlotID)[i] & fire_data$Timestep == "Pretreatment", "count.ha_Live_.all."]
}
## make mortality positive (i.e. = number of trees that die between pre and post7)
fire_data1_7$MortalityP_7 <- -1 * fire_data1_7$MortalityP_7

```


#### Fit models:

###### Total Fuels

```{r}
lmedata <- subset(fire_data1_7, fire_data1_7$Timestep == "Post_1")

options(na.action = na.fail)
lmetotal <- lmer(fuelchange_total ~ BAdif + BAdifSnag + SPPRatioABCO + SPPRatioCADE + SPPRatioOTHER + SPPRatioLIDE + SPPRatioQUKE + SPPRatioPSME + SPPRatioPIPO + difcons + (1|Unit), data = lmedata)

dredge(lmetotal)

lme1 <- lmer(fuelchange_total ~ BAdifSnag + SPPRatioABCO + SPPRatioCADE + SPPRatioPSME + SPPRatioQUKE + SPPRatioPIPO + (1|Unit), data = lmedata)

summary(lme1)
anova(lme1)
```


```{r}


lmetotalcount <- lmer(fuelchange_total ~ MortalityP_7 + count.ha_Live_ABCO + count.ha_Live_CADE + count.ha_Live_OTHER + count.ha_Live_LIDE + count.ha_Live_QUKE + count.ha_Live_PSME + count.ha_Live_PIPO + difcons + (1|Unit), data = lmedata)

dredge(lmetotalcount)

lme2 <- lmer(fuelchange_total ~ BAdifSnag + count.ha_Live_ABCO + count.ha_Live_CADE + count.ha_Live_PSME + count.ha_Live_QUKE + count.ha_Live_PIPO + (1|Unit), data = lmedata)

summary(lme2)
anova(lme2)

```

```{r}

lmemort <- lmer(fuelchange_total ~ MortalityP_7 +  + (1|Unit), data = lmedata)
summary(lmemort)
anova(lmemort)
```






```{r}

ggplot(lmedata, aes(MortalityP_7)) +
  geom_histogram(bins = 10)


ggplot(lmedata, aes(y = fuelchange_total, x = MortalityP_7)) +
  geom_point() + 
  geom_smooth()


```






