---
title: "FFSFuels_modeling"
author: "Jacob_Levine"
date: "9/18/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Run data cleaning script:
```{r}

source("FFSFuels_DataClean.R")

source("FFSFuels_EDA.Rmd")

```

**Jacob 9/25**
This is very incomplete, switched over to looking at consumption and accumulation in burn plots:



## what drives differences in rates of fuel accumulation 14 years after initial treatment?

### Prepare data

```{r}
## subset data
FFS_data_14 <- subset(FFS_data, TimestepNum == 14)

## calculate mean density (basal area):
meanBA <- mean(FFS_data_14$BA.mha.sum_Live_.all.)

## calculate mean density by block:
blockmeans <- aggregate(x = FFS_data$BA.mha.sum_Live_.all., by = list(FFS_data$Unit), FUN = mean)

## calculate difference between blockmeans and overall means
blockmeans$y <- blockmeans$x - meanBA

plotmeans <- aggregate(x = FFS_data$BA.mha.sum_Live_.all., by = list(FFS_data$Unit, FFS_data$PlotID), FUN = mean)

## append block means to FFS_data_14:
FFS_data_14$blockmean <- numeric(length = 207)
FFS_data_14$BAdif <- numeric(length = 207)
for (i in 1:nrow(blockmeans)) {
  FFS_data_14[FFS_data_14$Unit == blockmeans$Group.1[i], "blockmean"] <- blockmeans$x[i]
}

## so that BAdif is difference between average BA of plot for all times minus average BA of unit for all times:
for (i in 1:nrow(plotmeans)) {
  FFS_data_14[FFS_data_14$Unit == plotmeans$Group.1[i] &
                FFS_data_14$PlotID == plotmeans$Group.2[i], "BAdif"] <- plotmeans$x[i] - blockmeans[blockmeans$Group.1 == plotmeans$Group.1[i], "x"]
}

# for (i in 1:nrow(FFS_data_14)) {
#   for (j in 1:nrow(blockmeans)) {
#     if (FFS_data_14$Unit[i] == blockmeans$Group.1[j]) {
#       FFS_data_14$BAdif[i] <- FFS_data_14$BA.mha.sum_Live_.all.[i] - blockmeans$x[j]
#     }
#   }
# }

## make change the difference between fuelload at 14 and fuelload at 1:
FFS_data_1 <- subset(FFS_data, TimestepNum == 1)

```


### Fitting models

#### Total Fuels:

###### For change in fuelload7:
```{r}
## without density as cov
lme14ctotal <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_total ~ Treatment, 
               random = ~1|Unit)
summary(lme14ctotal)
anova(lme14ctotal)

## with density as cov:
lme14ctotal.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_total ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14ctotal.BA)
anova(lme14ctotal.BA)

## with density as cov and interaction:
lme14ctotal.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_total ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14ctotal.BAint)
anova(lme14ctotal.BAint)

## compare models:
AIC(lme14ctotal, lme14ctotal.BA, lme14ctotal.BAint)

```

###### For change in fuelload14:
```{r}
## without density as cov
lme14cctotal <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_total ~ Treatment, 
               random = ~1|Unit)
summary(lme14cctotal)
anova(lme14cctotal)

## with density as cov:
lme14cctotal.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_total ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14cctotal.BA)
anova(lme14cctotal.BA)

## with density as cov and interaction:
lme14cctotal.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_total ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14cctotal.BAint)
anova(lme14cctotal.BAint)

## compare models:
AIC(lme14cctotal, lme14cctotal.BA, lme14cctotal.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14rtotal <- lme(data = FFS_data_14, 
               fixed = fuelrate_total ~ Treatment, 
               random = ~1|Unit)
summary(lme14rtotal)
anova(lme14rtotal)

## with density as cov 
lme14rtotal.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_total ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14rtotal.BA)
anova(lme14rtotal.BA)

## with density as cov with interaction
lme14rtotal.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_total ~ Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14rtotal.BAint)
anova(lme14rtotal.BAint)

## compare models:
AIC(lme14rtotal, lme14rtotal.BA, lme14rtotal.BAint)

```

###### For stock:
```{r}
## without density as cov
lme14stotal <- lme(data = FFS_data_14, 
               fixed = logfuelload_total ~ Treatment, 
               random = ~1|Unit)
summary(lme14stotal)
anova(lme14stotal)

## with density as cov with interaction
lme14stotal.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_total ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14stotal.BA)
anova(lme14stotal.BA)

## with density as cov with interaction
lme14stotal.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_total ~ Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14stotal.BAint)
anova(lme14stotal.BAint)

## compare models
AIC(lme14stotal, lme14stotal.BA, lme14stotal.BAint)

```

For total fuel loads , best model is purely dependent on treatment.

**For change pre-7** 
Treatment effects significant for trt only model. Mechburn has greatest reduction, then burn, then mech.  Mech has almost no effect compared to control (p = .98). Note that on average the control experienced a reduction in this period of 11.6 tons/hectare.

**For change pre-14**
Treatment effects significant for best fitting model (BAint).  Mechburn has greatest reduction, then burn, then mech.  Mechburn and mech are comparable.

**For rates**
Main effect of mechanical treatment is largest (reduces the rate the most).  Followed by Burn, and then MechBurn.

**For Stocks**
Main effect of Burn is the largest followed by MechBurn and then Mech.


#### For Duff:

###### For change in fuelload:
```{r}
## without density as cov
lme14cduff <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_duff ~ Treatment, 
               random = ~1|Unit)
summary(lme14cduff)
anova(lme14cduff)

## with density as cov
lme14cduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_duff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14cduff.BA)
anova(lme14cduff.BA)

## with density as cov
lme14cduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_duff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14cduff.BAint)
anova(lme14cduff.BAint)

## compare models:
AIC(lme14cduff, lme14cduff.BA, lme14cduff.BAint)

```

###### For change in fuelload pre-14:
```{r}
## without density as cov
lme14ccduff <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_duff ~ Treatment, 
               random = ~1|Unit)
summary(lme14ccduff)
anova(lme14ccduff)

## with density as cov
lme14ccduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_duff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14ccduff.BA)
anova(lme14ccduff.BA)

## with density as cov
lme14ccduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_duff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14ccduff.BAint)
anova(lme14ccduff.BAint)

## compare models:
AIC(lme14ccduff, lme14cduff.BA, lme14cduff.BAint)

```


###### For rate of change:
```{r}
## without density as cov
lme14rduff <- lme(data = FFS_data_14, 
               fixed = fuelrate_duff ~ Treatment, 
               random = ~1|Unit)
summary(lme14rduff)
anova(lme14rduff)

## with density as cov
lme14rduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_duff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14rduff.BA)
anova(lme14rduff.BA)

## with density as cov with interactions
lme14rduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_duff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14rduff.BAint)
anova(lme14rduff.BAint)

## compare models:
AIC(lme14rduff, lme14rduff.BA, lme14rduff.BAint)

```

###### For stock:
```{r}
## without density as cov
lme14sduff <- lme(data = FFS_data_14, 
               fixed = logfuelload_duff ~ Treatment, 
               random = ~1|Unit)
summary(lme14sduff)
anova(lme14sduff)

## with density as cov
lme14sduff.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_duff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14sduff.BA)
anova(lme14sduff.BA)

## with density as cov with interactions
lme14sduff.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_duff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14sduff.BAint)
anova(lme14sduff.BAint)

## compare models
AIC(lme14sduff, lme14sduff.BA, lme14sduff.BAint)

```

**Net Change pre-7**
Best fitting model is just treatment.
Treatment effects are significant. Mechburn has most reduction, followed by burn

**Net Change pre-14**
Best fitting model includes density
Treatment effects and BA are both significant in terms of main effects.  BA has a negative relationship with duff fuelload.  Burn is strongest, followed by mechburn and then mech (not sig at p=0.05)

**Rate of Change**
Treatment effects not significant for best fitting model (just trt)

**LogStock**
Treatment effects significant for best fitting model (just trt). Burn decreases stock the most, then mechburn, then mech.  

```{r}

ggplot(FFS_data_14, aes(x = Treatment, y = logfuelload_duff)) + 
  geom_boxplot()

ggplot(FFS_data_14, aes(x = Treatment, y = fuelchangeBACI14_duff)) + 
  geom_boxplot()

ggplot(FFS_data_14, aes(x = BAdif, y = fuelchangeBACI14_duff)) + 
  geom_point() + 
  geom_smooth(size = 0.5) 




```


#### For Litter:

###### For change in fuelload:
```{r}
## without density as cov
lme14clitter <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litter ~ Treatment, 
               random = ~1|Unit)
summary(lme14clitter)
anova(lme14clitter)

## with density as cov
lme14clitter.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litter ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14clitter.BA)
anova(lme14clitter.BA)

## with density as cov
lme14clitter.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litter ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14clitter.BAint)
anova(lme14clitter.BAint)

## compare models:
AIC(lme14clitter, lme14clitter.BA, lme14clitter.BAint)

```


###### For change in fuelload pre-14:
```{r}
## without density as cov
lme14cclitter <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litter ~ Treatment, 
               random = ~1|Unit)
summary(lme14cclitter)
anova(lme14cclitter)

## with density as cov
lme14cclitter.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litter ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14cclitter.BA)
anova(lme14cclitter.BA)

## with density as cov
lme14cclitter.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litter ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14cclitter.BAint)
anova(lme14cclitter.BAint)

## compare models:
AIC(lme14cclitter, lme14cclitter.BA, lme14cclitter.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14rlitter <- lme(data = FFS_data_14, 
               fixed = fuelrate_litter ~ Treatment, 
               random = ~1|Unit)
summary(lme14rlitter)
anova(lme14rlitter)

## with density as cov
lme14rlitter.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_litter ~ Treatment + BAdif , 
               random = ~1|Unit)

summary(lme14rlitter.BA)
anova(lme14rlitter.BA)

## with density as cov and interactions
lme14rlitter.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_litter ~ Treatment * BAdif , 
               random = ~1|Unit)

summary(lme14rlitter.BAint)
anova(lme14rlitter.BAint)

## compare models:
AIC(lme14rlitter, lme14rlitter.BA, lme14rlitter.BAint)

```

###### For stock:
```{r}
## without density as cov
lme14slitter <- lme(data = FFS_data_14, 
               fixed = logfuelload_litter ~ Treatment, 
               random = ~1|Unit)
summary(lme14slitter)
anova(lme14slitter)

## with density as cov
lme14slitter.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_litter ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14slitter.BA)
anova(lme14slitter.BA)

## with density as cov with interactions
lme14slitter.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_litter ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14slitter.BAint)
anova(lme14slitter.BAint)

## compare models
AIC(lme14slitter, lme14slitter.BA, lme14slitter.BAint)

```

**For change pre-7**
Treatment effects not significant for best fitting model (int only).

**For change pre-14**
Treatment effects not significant for best fitting model (int only)

**For rate of change**
Treatment effects are not significant for best fitting model (int only).

**For Stock**
Treatment effects are not significant for best fitting model (int only).


#### For Litterduff:

###### For change in logfuelload:
```{r}
## without density as cov
lme14clitterduff <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litterduff ~ Treatment, 
               random = ~1|Unit)
summary(lme14clitterduff)
anova(lme14clitterduff)

## with density as cov
lme14clitterduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litterduff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14clitterduff.BA)
anova(lme14clitterduff.BA)

## with density as cov
lme14clitterduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_litterduff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14clitterduff.BAint)
anova(lme14clitterduff.BAint)

## compare models:
AIC(lme14clitterduff, lme14clitterduff.BA, lme14clitterduff.BAint)

```

###### For change in logfuelload pre-14:
```{r}
## without density as cov
lme14cclitterduff <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litterduff ~ Treatment, 
               random = ~1|Unit)
summary(lme14cclitterduff)
anova(lme14cclitterduff)

## with density as cov
lme14cclitterduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litterduff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14cclitterduff.BA)
anova(lme14cclitterduff.BA)

## with density as cov
lme14cclitterduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_litterduff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14cclitterduff.BAint)
anova(lme14cclitterduff.BAint)

## compare models:
AIC(lme14cclitterduff, lme14cclitterduff.BA, lme14cclitterduff.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14rlitterduff <- lme(data = FFS_data_14, 
               fixed = fuelrate_litterduff ~ Treatment, 
               random = ~1|Unit)
summary(lme14rlitterduff)
anova(lme14rlitterduff)

## with density as cov
lme14rlitterduff.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_litterduff ~ Treatment + BAdif, 
               random = ~1|Unit)

## with density as cov and interactions
lme14rlitterduff.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_litterduff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14rlitterduff.BAint)
anova(lme14rlitterduff.BAint)

## compare models:
AIC(lme14rlitterduff, lme14rlitterduff.BA, lme14rlitterduff.BAint)

```

###### For Logstock:
```{r}
## without density as cov
lme14slitterduff <- lme(data = FFS_data_14, 
               fixed = logfuelload_litterduff~ Treatment, 
               random = ~1|Unit)
summary(lme14slitterduff)
anova(lme14slitterduff)

## with density as cov
lme14slitterduff.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_litterduff ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14slitterduff.BA)
anova(lme14slitterduff.BA)

## with density as cov
lme14slitterduff.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_litterduff ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14slitterduff.BAint)
anova(lme14slitterduff.BAint)

## compare models
AIC(lme14slitterduff, lme14slitterduff.BA, lme14slitterduff.BAint)

```

**For Stock pre-7**
Treatment effects are significant for best model (trt only).  Mechburn > Burn > Mech (not significant).

**For Stock pre-14**
Treatment effects and density significant for best model (BAint).  Mechburn > Burn > Mech (not significant).  Again, BA has a negative relationship with fuelload (however this could be difficult to interpret given interactions).  Interactions are not significant.

**For Rate** 
Treatment effects not significant.

**For stock**
Treatment effects are significant, Burn reduces the most, followed by mechburn and then mech.




#### For 1h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c1h <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_1h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c1h)
anova(lme14c1h)

## with density as cov
lme14c1h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14c1h.BA)
anova(lme14c1h.BA)

## with density as cov with interaction
lme14c1h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_1h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14c1h.BAint)
anova(lme14c1h.BAint)

## compare models:
AIC(lme14c1h, lme14c1h.BA, lme14c1h.BAint)

```

###### For change in logfuelload:
```{r}
## without density as cov
lme14cc1h <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_1h ~ Treatment, 
               random = ~1|Unit)
summary(lme14cc1h)
anova(lme14cc1h)

## with density as cov
lme14cc1h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14cc1h.BA)
anova(lme14cc1h.BA)

## with density as cov with interaction
lme14cc1h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI14_1h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14cc1h.BAint)
anova(lme14cc1h.BAint)

## compare models:
AIC(lme14cc1h, lme14cc1h.BA, lme14cc1h.BAint)

```

###### For change in logfuelload:
```{r}
## without density as cov
lme14c1h <- lme(data = FFS_data_14, 
               fixed = fuelchange_1h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c1h)
anova(lme14c1h)

## with density as cov
lme14c1h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14c1h.BA)
anova(lme14c1h.BA)

## with density as cov with interaction
lme14c1h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchange_1h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14c1h.BAint)
anova(lme14c1h.BAint)

## compare models:
AIC(lme14c1h, lme14c1h.BA, lme14c1h.BAint)

```


###### For rate of change:
```{r}
## without density as cov
lme14r1h <- lme(data = FFS_data_14, 
               fixed = fuelrate_1h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r1h)
anova(lme14r1h)

## with density as cov
lme14r1h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14r1h.BA)
anova(lme14r1h.BA)

## with density as cov
lme14r1h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_1h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14r1h.BAint)
anova(lme14r1h.BAint)

## compare models:
AIC(lme14r1h, lme14r1h.BA, lme14r1h.BAint)

```

###### For stock:

Issue of zero inflation comes up here 
```{r}
## without density as cov
lme14s1h <- lme(data = FFS_data_14, 
               fixed = logfuelload_1h ~ Treatment, 
               random = ~1|Unit)
summary(lme14s1h)
anova(lme14s1h)

## with density as cov
lme14s1h.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14s1h.BA)
anova(lme14s1h.BA)

## with density as cov with interactions
lme14s1h.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_1h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14s1h.BAint)
anova(lme14s1h.BAint)

## compare models
AIC(lme14s1h, lme14s1h.BA, lme14s1h.BAint)

```

**For change**
Treatment effects not significant.

**For change pre-14**

**For rate of change**
Treatment effects not significant.

**For stock**
Treatment effects not significant at 0.05 level, but clost (p = 0.066), interesting that in this case mech has strongest effect on stock, followed by mechburn and then mech


#### For 10h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c10h <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_10h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c10h)
anova(lme14c10h)

## with density as cov
lme14c10h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_10h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14c10h.BA)
anova(lme14c10h.BA)

## with density as cov
lme14c10h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchangeBACI_10h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14c10h.BAint)
anova(lme14c10h.BAint)

## compare models:
AIC(lme14c10h, lme14c10h.BA, lme14c10h.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14r10h <- lme(data = FFS_data_14, 
               fixed = fuelrate_10h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r10h)
anova(lme14r10h)

## with density as cov
lme14r10h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_10h ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14r10h.BA)
anova(lme14r10h.BA)

## compare models:
AIC(lme14r10h, lme14r10h.BA)

```

###### For stock:
```{r}
## without density as cov
lme14s10h <- lme(data = FFS_data_14, 
               fixed = logfuelload_10h_tonha ~ Treatment, 
               random = ~1|Unit)
summary(lme14s10h)
anova(lme14s10h)

## with density as cov
lme14s10h.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_10h_tonha ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14s10h.BA)
anova(lme14s10h.BA)

## compare models
AIC(lme14s10h, lme14s10h.BA)

```
**For change pre-7**
Treatment effects not significant.
**For change pre-14**
Treatment effects not significant.
**For rate of change**

#### For 100h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c100h <- lme(data = FFS_data_14, 
               fixed = fuelchange_100h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c100h)
anova(lme14c100h)

## with density as cov
lme14c100h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_100h ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14c100h.BA)
anova(lme14c100h.BA)

## compare models:
AIC(lme14c100h, lme14c100h.BA)

```

###### For rate of change:
```{r}
## without density as cov
lme14r100h <- lme(data = FFS_data_14, 
               fixed = fuelrate_100h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r100h)
anova(lme14r100h)

## with density as cov
lme14r100h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_100h ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14r100h.BA)
anova(lme14r100h.BA)

## compare models:
AIC(lme14r100h, lme14r100h.BA)

```

###### For stock:

Deal with zero-inflation by using a ZIP/ZINB model:
```{r}
library(pscl)

f1 <- formula(logfuelload_100h_tonha ~ Treatment)
zilm100h <- zeroinfl(f1, data = FFS_data_14, dist= "negbin")


## without density as cov
lme14s100h <- lme(data = FFS_data_14, 
               fixed = logfuelload_100h_tonha ~ Treatment, 
               random = ~1|Unit)
summary(lme14s100h)
anova(lme14s100h)

lm14s100h <- lm(data = FFS_data_14, 
               logfuelload_100h_tonha ~ Treatment)
summary(lm14s100h)
anova(lm14s100h)


## with density as cov
lme14s100h.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_100h_tonha ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14s100h.BA)
anova(lme14s100h.BA)

## compare models
AIC(lme14s100h, lme14s100h.BA)

```


#### For 1000h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c1000h <- lme(data = FFS_data_14, 
               fixed = fuelchange_1000h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c1000h)
anova(lme14c1000h)

## with density as cov
lme14c1000h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_1000h ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14c1000h.BA)
anova(lme14c1000h.BA)

## compare models:
AIC(lme14c1000h, lme14c1000h.BA)

```

###### For rate of change:
```{r}
## without density as cov
lme14r1000h <- lme(data = FFS_data_14, 
               fixed = fuelrate_1000h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r1000h)
anova(lme14r1000h)

## with density as cov
lme14r1000h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_1000h ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14r1000h.BA)
anova(lme14r1000h.BA)

## compare models:
AIC(lme14r1000h, lme14r1000h.BA)

```

###### For stock:

Dealing with zero-inflation - fit a two-step hurdle model
```{r}
## First step: is there at least one 1000hr present:
## 1) create binomial variable for detection/non-detection
FFS_data_14$bin1000h <- rep(NA, times = nrow(FFS_data_14))
for (i in 1:nrow(FFS_data_14)) {
  if (FFS_data_14$logfuelload_1000h_tonha[i] == 0) {
    FFS_data_14$bin1000h[i] <- 0
  }
  else if (FFS_data_14$logfuelload_1000h_tonha[i] > 0) {
    FFS_data_14$bin1000h[i] <- 1
  }
}
## compare mixed effects to fixed effects model
lme14s1000hpresM <- glmer(formula = bin1000h ~ 1|Unit, data = FFS_data_14, family = binomial)
summary(lme14s1000hpresM)

lme14s1000hpresF <- glm(formula = bin1000h ~ 1, data = FFS_data_14, family = binomial)
summary(lme14s1000hpresF)
## fixed effects model minimizes AIC

## 2) remove 0s and fit lme model: 
for (i in 1:nrow(FFS_data_14)) {
  if (FFS_data_14$logfuelload_1000h_tonha[i] == 0) {
    FFS_data_14$logfuelload_1000h_tonha[i] <- NA
  }
}

FFS_data_14_sub <- FFS_data_14[complete.cases(FFS_data_14$logfuelload_1000h_tonha),]

## without density as cov
lme14s1000h <- lme(data = FFS_data_14_sub, 
               fixed = logfuelload_1000h_tonha ~ Treatment, 
               random = ~1|Unit)
summary(lme14s1000h)
anova(lme14s1000h)

## with density as cov
lme14s1000h.BA <- lme(data = FFS_data_14_sub, 
               fixed = logfuelload_1000h_tonha ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14s1000h.BA)
anova(lme14s1000h.BA)

## compare models
AIC(lme14s1000h, lme14s1000h.BA)

```


#### For 1.100h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c1.100h <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.100h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c1.100h)
anova(lme14c1.100h)

## with density as cov
lme14c1.100h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.100h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14c1.100h.BA)
anova(lme14c1.100h.BA)

## with density as cov
lme14c1.100h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.100h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14c1.100h.BAint)
anova(lme14c1.100h.BAint)

## compare models:
AIC(lme14c1.100h, lme14c1.100h.BA, lme14c1.100h.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14r1.100h <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.100h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r1.100h)
anova(lme14r1.100h)

## with density as cov
lme14r1.100h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.100h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14r1.100h.BA)
anova(lme14r1.100h.BA)

## with density as cov
lme14r1.100h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.100h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14r1.100h.BAint)
anova(lme14r1.100h.BAint)

## compare models:
AIC(lme14r1.100h, lme14r1.100h.BA, lme14r1.100h.BAint)

```

###### For log of stock:
```{r}
## without density as cov
lme14s1.100h <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.100h ~ Treatment, 
               random = ~1|Unit)
summary(lme14s1.100h)
anova(lme14s1.100h)

## with density as cov
lme14s1.100h.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.100h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14s1.100h.BA)
anova(lme14s1.100h.BA)

## with density as cov
lme14s1.100h.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.100h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14s1.100h.BAint)
anova(lme14s1.100h.BAint)

## compare models
AIC(lme14s1.100h, lme14s1.100h.BA, lme14s1.100h.BAint)

```
**For change**
Treatment effects not significant.

**For rate of change**
Treatment effects not significant.

**For log of stock**
Treatment effects not significant.


#### For 1.1000h:

###### For change in logfuelload:
```{r}
## without density as cov
lme14c1.1000h <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.1000h ~ Treatment, 
               random = ~1|Unit)
summary(lme14c1.1000h)
anova(lme14c1.1000h)

## with density as cov
lme14c1.1000h.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.1000h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14c1.1000h.BA)
anova(lme14c1.1000h.BA)

## with density as cov with interactions
lme14c1.1000h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelchange_1.1000h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14c1.1000h.BAint)
anova(lme14c1.1000h.BAint)

## compare models:
AIC(lme14c1.1000h, lme14c1.1000h.BA, lme14c1.1000h.BAint)

```

###### For rate of change:
```{r}
## without density as cov
lme14r1.1000h <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.1000h ~ Treatment, 
               random = ~1|Unit)
summary(lme14r1.1000h)
anova(lme14r1.1000h)

## with density as cov
lme14r1.1000h.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.1000h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14r1.1000h.BA)
anova(lme14r1.1000h.BA)

## with density as cov
lme14r1.1000h.BAint <- lme(data = FFS_data_14, 
               fixed = fuelrate_1.1000h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14r1.1000h.BAint)
anova(lme14r1.1000h.BAint)

## compare models:
AIC(lme14r1.1000h, lme14r1.1000h.BA, lme14r1.1000h.BAint)

```

###### For stock:
```{r}
## without density as cov
lme14s1.1000h <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.1000h ~ Treatment, 
               random = ~1|Unit)
summary(lme14s1.1000h)
anova(lme14s1.1000h)

## with density as cov
lme14s1.1000h.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.1000h ~ Treatment + BAdif, 
               random = ~1|Unit)

summary(lme14s1.1000h.BA)
anova(lme14s1.1000h.BA)

## with density as cov
lme14s1.1000h.BAint <- lme(data = FFS_data_14, 
               fixed = logfuelload_1.1000h ~ Treatment * BAdif, 
               random = ~1|Unit)

summary(lme14s1.1000h.BAint)
anova(lme14s1.1000h.BAint)

## compare models
AIC(lme14s1.1000h, lme14s1.1000h.BA, lme14s1.1000h.BAint)

```
Treatment effects not significant for any of the cases.


#### For surface:

###### For change in logfuelload:
```{r}
## without density as cov
lme14csurface <- lme(data = FFS_data_14, 
               fixed = fuelchange_surface ~ Treatment, 
               random = ~1|Unit)
summary(lme14csurface)
anova(lme14csurface)

## with density as cov
lme14csurface.BA <- lme(data = FFS_data_14, 
               fixed = fuelchange_surface ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14csurface.BA)
anova(lme14csurface.BA)

## compare models:
AIC(lme14csurface, lme14csurface.BA)

```

###### For rate of change:
```{r}
## without density as cov
lme14rsurface <- lme(data = FFS_data_14, 
               fixed = fuelrate_surface ~ Treatment, 
               random = ~1|Unit)
summary(lme14rsurface)
anova(lme14rsurface)

## with density as cov
lme14rsurface.BA <- lme(data = FFS_data_14, 
               fixed = fuelrate_surface ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14rsurface.BA)
anova(lme14rsurface.BA)

## compare models:
AIC(lme14rsurface, lme14rsurface.BA)

```

###### For stock:
```{r}
## without density as cov
lme14ssurface <- lme(data = FFS_data_14, 
               fixed = logfuelload_surface_tonha ~ Treatment, 
               random = ~1|Unit)
summary(lme14ssurface)
anova(lme14ssurface)

## with density as cov
lme14ssurface.BA <- lme(data = FFS_data_14, 
               fixed = logfuelload_surface_tonha ~ Treatment + BAdif + Treatment*BAdif, 
               random = ~1|Unit)

summary(lme14ssurface.BA)
anova(lme14ssurface.BA)

## compare models
AIC(lme14ssurface, lme14ssurface.BA)

```






##**Including Timestep as a determinant**

###### For change in logfuelload:

```{r}
## for change:
lmeTctotal <- lme(data = subset(FFS_data, FFS_data$TimestepNum != -1), 
                fixed = fuelchange_total ~ Treatment + Timestep + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTctotal)
anova(lmeTctotal)

```

###### For rate:
```{r}
## for change:
lmeTrtotal <- lme(data = subset(FFS_data, FFS_data$TimestepNum != -1), 
                fixed = fuelrate_total ~ Treatment + Timestep + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTrtotal)
anova(lmeTrtotal)

lmeTrtotalBA <- lme(data = subset(FFS_data, FFS_data$TimestepNum != -1), 
                fixed = fuelrate_total ~ Treatment + Timestep + BA.mha.sum_Live_.all. + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTrtotalBA)
anova(lmeTrtotalBA)

AIC(lmeTrtotal, lmeTrtotalBA)
```

###### For stock:
```{r}

lmeTstotal <- lme(data = FFS_data, 
                fixed = logfuelload_total_tonha ~ Treatment + Timestep + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTstotal)
anova(lmeTstotal)

lmeTstotalBA <- lme(data = FFS_data, 
                fixed = logfuelload_total_tonha ~ Treatment + Timestep + BA.mha.sum_Live_.all. + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTstotalBA)
anova(lmeTstotalBA)

AIC(lmeTstotal, lmeTstotalBA)

```


```{r}
## for change:
lmeTrtotal <- lme(data = subset(FFS_data, FFS_data$TimestepNum == 1 | FFS_data$TimestepNum == 7), 
                fixed = fuelrate_total ~ Treatment + Timestep + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTrtotal)
anova(lmeTrtotal)

lmeTrtotalBA <- lme(data = subset(FFS_data, FFS_data$TimestepNum != -1), 
                fixed = fuelrate_total ~ Treatment + Timestep + BA.mha.sum_Live_.all. + Treatment*Timestep, 
                random = list(~1|Unit, ~1|PlotID))

summary(lmeTrtotalBA)
anova(lmeTrtotalBA)

AIC(lmeTrtotal, lmeTrtotalBA)


```






